### 2024-06-15 code/eda-code/EDA-Code.qmd
Completed Final Project - Part 2
Rewrote/reformatted some wording in the manuscript. Cleaned and explored the data, creating plots of distributions and correlations, and then added them to the manuscript. Removed files that did not go with my project/from the template. Made sure it was reproducible and was able to knit into an html.

---
title: "EDA Code"
subtitle: ""
author: Austin Thrash
date: today
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
    highlight-style: github
csl: ../../assets/apa.csl
---

```{r}
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(readr)
library(tidyr) # data wrangling
library(rpart) # DT
library(tidyverse)
library(car)
library(tree)
library(caret)
library(MASS)
library(ROCR)
library(randomForest)
library(corrplot)
library(ggcorrplot) # corr plot
library(reshape2) # heat map
library(GGally) # Pais ggplot2
```

# Reading in the Data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Simple Exploration

```{r}

```

# Plots of Response/Features

## Histogram

```{r}
p1 <- mydata %>% ggplot(aes(x=Diabetes_012)) + geom_histogram() 
plot(p1)
figure_file = here("results", "figures","diabetes-distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

```{r}
p2 <- mydata %>% ggplot(aes(x=Smoker)) + geom_histogram() 
plot(p2)
figure_file = here("results", "figures","smoker-distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

```{r}
p3 <- mydata %>% ggplot(aes(x=BMI)) + geom_histogram() 
plot(p3)
figure_file = here("results", "figures","BMI-distribution.png")
ggsave(filename = figure_file, plot=p3) 
```

```{r}
p4 <- mydata %>% ggplot(aes(x=PhysHlth)) + geom_histogram() 
plot(p4)
figure_file = here("results", "figures","PhysHlth-distribution.png")
ggsave(filename = figure_file, plot=p4) 
```

```{r}
p5 <- mydata %>% ggplot(aes(x=Age)) + geom_histogram() 
plot(p5)
figure_file = here("results", "figures","Age-distribution.png")
ggsave(filename = figure_file, plot=p5) 
```

## Correlation Matrix

```{r}
library(corrplot)

# Compute the correlation matrix for numerical attributes
numerical_data <- mydata[, sapply(mydata, is.numeric)]
cor_matrix <- cor(numerical_data)  # Handles missing values by excluding them

# Define the file path to save the plot
figure_file <- here("results", "figures", "correlation_matrix.png")

# Save the correlation matrix plot
png(filename = figure_file, width = 800, height = 600)
corrplot(cor_matrix, method = "color", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, addCoef.col = "black", number.cex = 0.5)
dev.off()
```

```{r}
mydata$Diabetes_012 <- as.factor(mydata$Diabetes_012)
```

# Features of Importance

## Linear Model

```{r}
glm.full = glm(Diabetes_012 ~ ., data = mydata, family = binomial)

glm.vif = vif(glm.full)

glm.vif
```

## Random Forest

```{r}
set.seed(12345)
import_features = train(Diabetes_012 ~ ., 
               data = mydata, 
               method = "rf", 
               importance = TRUE, 
               trControl = trainControl(method = "cv", number = 2, verbose = TRUE))
```

```{r}
varimp_plot <- plot(varImp(import_features))

# Define the file path to save the plot
figure_file <- here("results", "figures", "variable_importance.png")

# Save the plot
png(filename = figure_file, width = 1000, height = 800)
print(varimp_plot)
dev.off()
```