### 2024-06-22 code/eda-code/EDA-Code.qmd

---
title: "EDA Code"
subtitle: ""
author: Austin Thrash
date: today
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
    highlight-style: github
csl: ../../assets/apa.csl
---

```{r}
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(readr)
library(tidyr) # data wrangling
library(rpart) # DT
library(tidyverse)
library(car)
library(tree)
library(caret)
library(MASS)
library(ROCR)
library(randomForest)
library(corrplot)
library(ggcorrplot) # corr plot
library(reshape2) # heat map
library(GGally) # Pais ggplot2
library(gridExtra)
```

# Reading in the Data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Simple Exploration

# Plots of Response/Features

## Histogram

```{r}
# Function to create histograms for each variable
create_histograms <- function(data) {
  plots <- list()
  for (col_name in names(data)) {
    if (is.numeric(data[[col_name]])) {
      p <- ggplot(data, aes_string(x = col_name)) + 
        geom_histogram(binwidth = 1, fill = "salmon", color = "black") +
        labs(title = col_name) +
        theme_minimal()
    } else {
      p <- ggplot(data, aes_string(x = col_name)) + 
        geom_bar(fill = "salmon", color = "black") +
        labs(title = col_name) +
        theme_minimal()
    }
    plots[[col_name]] <- p
  }
  return(plots)
}

# Create histograms for all variables
histograms <- create_histograms(mydata)

# Arrange the plots in a grid
n <- length(histograms)
nCol <- ceiling(sqrt(n))
nRow <- ceiling(n / nCol)
combined_plot <- marrangeGrob(grobs = histograms, nrow = nRow, ncol = nCol)

# Save the combined plot to a PNG file
figure_file <- here("results", "figures", "variable-distributions.png")
ggsave(filename = figure_file, plot = combined_plot, width = 20, height = 20)
```

## Correlation Matrix

```{r}
library(corrplot)

# Compute the correlation matrix for numerical attributes
numerical_data <- mydata[, sapply(mydata, is.numeric)]
cor_matrix <- cor(numerical_data)  # Handles missing values by excluding them

# Define the file path to save the plot
figure_file <- here("results", "figures", "correlation_matrix.png")

# Save the correlation matrix plot
png(filename = figure_file, width = 800, height = 600)
corrplot(cor_matrix, method = "color", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, addCoef.col = "black", number.cex = 0.5)
dev.off()
```

```{r}
mydata$Diabetes_012 <- as.factor(mydata$Diabetes_012)
```

# Features of Importance

## Linear Model

```{r}
glm.full = glm(Diabetes_012 ~ ., data = mydata, family = binomial)

glm.vif = vif(glm.full)

glm.vif
```

## Random Forest

```{r}
set.seed(12345)
import_features = train(Diabetes_012 ~ ., 
               data = mydata, 
               method = "rf", 
               importance = TRUE, 
               trControl = trainControl(method = "cv", number = 2, verbose = TRUE))
```

```{r}
varimp_plot <- plot(varImp(import_features))

# Define the file path to save the plot
figure_file <- here("results", "figures", "variable_importance.png")

# Save the plot
png(filename = figure_file, width = 1000, height = 800)
print(varimp_plot)
dev.off()
```