### 2024-06-22 products/manuscript/Manuscript.qmd

---
title: "Manuscript for a Data Analysis Project: Diabetes Diagnosis"
subtitle: ""
author: Austin Thrash
date: today
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
    highlight-style: github
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

*\*Summary will be finalized in later part of project and will include setup and results*

{{< pagebreak >}}

# Introduction

## General Background Information

*\*more background information will be added later, I will describe what diabetes is, what are the notable influences on diagnosis, and what the dataset provides in relation to other studies on diabetes diagnosis.*

## Description of data and data source

This dataset comes from UC Irvine Machine Learning Repository. This data has been gathered and processed from the CDCs website in 2014. This dataset includes over 250,000 instances (observations) and 21 features, ranging from personal information (income/education) to health history (drinking/smoking/physical health). Bellow are the features and there respective descriptions

| Feature              | Description                                                                                                                                                                                                                                                                                                                             |
|------------------|------------------------------------------------------|
| ID                   | Patient ID                                                                                                                                                                                                                                                                                                                              |
| Diabetes_binary      | 0 = no diabetes 1 = prediabetes or diabetes                                                                                                                                                                                                                                                                                             |
| HighBP               | 0 = no high BP 1 = high BP                                                                                                                                                                                                                                                                                                              |
| HighChol             | 0 = no high cholesterol 1 = high cholesterol                                                                                                                                                                                                                                                                                            |
| CholCheck            | 0 = no cholesterol check in 5 years 1 = yes cholesterol check in 5 years                                                                                                                                                                                                                                                                |
| BMI                  | Body Mass Index                                                                                                                                                                                                                                                                                                                         |
| Smoker               | Have you smoked at least 100 cigarettes in your entire life? \[Note: 5 packs = 100 cigarettes\] 0 = no 1 = yes                                                                                                                                                                                                                          |
| Stoke                | (Ever told) you had a stroke. 0 = no 1 = yes                                                                                                                                                                                                                                                                                            |
| HeartDiseaseorAttack | coronary heart disease (CHD) or myocardial infarction (MI) 0 = no 1 = yes                                                                                                                                                                                                                                                               |
| PhysActivity         | physical activity in past 30 days - not including job 0 = no 1 = yes                                                                                                                                                                                                                                                                    |
| Fruits               | Consume Fruit 1 or more times per day 0 = no 1 = yes                                                                                                                                                                                                                                                                                    |
| Veggies              | Consume Vegetables 1 or more times per day 0 = no 1 = yes                                                                                                                                                                                                                                                                               |
| HvyAlchoholConsump   | Heavy drinkers (adult men having more than 14 drinks per week and adult women having more than 7 drinks per week) 0 = no 1 = yes                                                                                                                                                                                                        |
| AnyHealthcare        | Have any kind of health care coverage, including health insurance, prepaid plans such as HMO, etc. 0 = no 1 = yes                                                                                                                                                                                                                       |
| NoDocbcCost          | Was there a time in the past 12 months when you needed to see a doctor but could not because of cost? 0 = no 1 = yes                                                                                                                                                                                                                    |
| GenHlth              | Would you say that in general your health is: scale 1-5 1 = excellent 2 = very good 3 = good 4 = fair 5 = poor                                                                                                                                                                                                                          |
| MentHlth             | Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good? scale 1-30 days                                                                                                                                        |
| PhysHlth             | Now thinking about your physical health, which includes physical illness and injury, for how many days during the past 30 days was your physical health not good? scale 1-30 days                                                                                                                                                       |
| DiffWalk             | Do you have serious difficulty walking or climbing stairs? 0 = no 1 = yes                                                                                                                                                                                                                                                               |
| Sex                  | 0 = female 1 = male                                                                                                                                                                                                                                                                                                                     |
| Age                  | 13-level age category (\_AGEG5YR see codebook) 1 = 18-24 9 = 60-64 13 = 80 or older                                                                                                                                                                                                                                                     |
| Education            | Education level (EDUCA see codebook) scale 1-6 1 = Never attended school or only kindergarten 2 = Grades 1 through 8 (Elementary) 3 = Grades 9 through 11 (Some high school) 4 = Grade 12 or GED (High school graduate) 5 = College 1 year to 3 years (Some college or technical school) 6 = College 4 years or more (College graduate) |
| Income               | Income scale (INCOME2 see codebook) scale 1-8 1 = less than \$10,000 5 = less than \$35,000 8 = \$75,000 or more                                                                                                                                                                                                                        |

<https://archive.ics.uci.edu/dataset/891/cdc+diabetes+health+indicators>

## Questions/Hypotheses to be addressed

"What are the key health indicators that predict the likelihood of an individual being diagnosed with diabetes?" Specifically, identifying which factors are most strongly associated with diabetes prevalence in the population.

{{< pagebreak >}}

# Methods

*\*Analysis methods will be included in later part of project*

```{r}
library(readr)
```

## Data acquisition

### Explain the question to be answered using the data

The main question I would like to answer with this project is "What are the key health indicators that predict the likelihood of an individual being diagnosed with diabetes?".

### What will be your outcome(s) of interest (if any)?

The primary outcome of interest will be whether an individual has been diagnosed with diabetes or not. This outcome will be represented as a binary variable where 1 indicates a diabetes diagnosis and 0 indicates no diabetes.

### What (if any) specific predictors will you focus on?

The specific predictors I will focus on are age, gender, BMI, consumption of fruits and vegetables, and frequency of exercise. Although these are the main predictors I will focus on, I'll also use data analytic techniques to identify which predictors are of importance or have relationships with the response variable.

### What relations/patterns are you looking for in the data?

I will look for patterns ad relationships between the predictors and the outcome of diabetes diagnosis. More specifically, I will analyze correlations between individual predictors and the response variable, interaction effects between predictors and the response variable, and predictive models such as logistic regression, decision trees, and machine learning classifiers.

## Data import and cleaning

Not much occurs during the cleaning stage as the dataset was processed and cleaned prior. The only modifications made to the dataset was removing the 'ID' column as it is not useful for this analysis. Once the data was imported, it was only a matter of converting categorical variables to factors so they would function properly. One-hot encoding was implemented, but only for the multinomial logistic regression.

## Statistical analysis method

For my analysis, I will first look at the distributions of each variable to identify if the dataset is balanced and outliers. I do not think any data cleaning will need to be done as that was completed when the data was collected from the CDCs website. Next, I will create a correlation matrix to view the correlations between each predictor and the response variable. After viewing the correlations, I will then use simple linear regression and random forest models to identify variables of importance, these variables will be noted and used later during the modeling phase. I will utilize cross validation during the modeling phase. This section will be filled out more once we are further into the class and have discussed more approaches/methods.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

### Distribution of all variables

```{r}
knitr::include_graphics(here("results","figures","variable-distributions.png"))
```

### Correlation Matrix

```{r}
knitr::include_graphics(here("results","figures","correlation_matrix.png"))
```

### Feature Importance - Linear Regression

\*Will fix and put into readable table

```         
            HighBP             HighChol            CholCheck                  BMI               Smoker               Stroke HeartDiseaseorAttack 
            1.141777             1.070725             1.011236             1.124242             1.073702             1.071638             1.150480 
        PhysActivity               Fruits              Veggies    HvyAlcoholConsump        AnyHealthcare          NoDocbcCost              GenHlth 
            1.140832             1.109908             1.105040             1.013591             1.100212             1.157580             1.743109 
            MentHlth             PhysHlth             DiffWalk                  Sex                  Age            Education               Income 
            1.281663             1.697022             1.499052             1.109269             1.326329             1.299898             1.505502 
```

### Feature Importance - Random Forest

```{r}
knitr::include_graphics(here("results","figures","variable_importance.png"))
```

Overall, the variables do not show much correlation with one another, however I have a feeling that I may need to do more data processing, i.e. making categorical variables into factors and removing them from the correlation matrix. I decided to include some distributions of variables that I found to be interesting, such as BMI, Age, PhysHlth, and Smoker. After running both a linear and random forest model to determine the most important features, it looks as if BMI, HighBP, HeartDiseaseorAttack, and HvyAlchoholConsump are the most important features to include in the models.

## Basic statistical analysis

***\*IN PROGRESS***

### Methods

After processing and exploring the data, it was discovered that a majority of the features in this dataset are categorical. Therefore, we will be making use of models that are more rigorous to our situation, such as random forest and KNN. However before trying those models, we will first do a simple multinomial logistic regression. Why multinomial? That is because our response variable is not binary, it has three levels to it; typically, logistic regression is used for binary classification. After creating models with all features included, we will go on to look at models with features that fit our importance criteria (i.e. look at the models results when using the most important features).

### Preparing the data

For some of the models, mainly multinomial logistic regression, we need to further process our data. The method we will use is known as one-hot encoding. One-hot encoding is a process of converting categorical variables into a binary (0 or 1) format where each category becomes a new binary column. We need to do this because most machine learning algorithms require numerical input and cannot handle categorical data directly, so formatting this data in a different manner will provide us with better results.

### Results

#### Multinomial Logistic Regression - All Features

```         
Confusion Matrix and Statistics

          Reference
Prediction    X0    X1    X2
        X0 21075   398  2891
        X1     3     1     1
        X2   477    32   622

Overall Statistics

               Accuracy : 0.8509          
                 95% CI : (0.8465, 0.8553)
    No Information Rate : 0.8453          
    P-Value [Acc > NIR] : 0.006606

                  Kappa : 0.1995

 Mcnemar's Test P-Value : < 2.2e-16

Statistics by Class:

                     Class: X0 Class: X1 Class: X2
Sensitivity             0.9777 2.320e-03   0.17701
Specificity             0.1663 9.998e-01   0.97685
Pos Pred Value          0.8650 2.000e-01   0.54996
Neg Pred Value          0.5775 9.831e-01   0.88132
Prevalence              0.8453 1.690e-02   0.13780
Detection Rate          0.8265 3.922e-05   0.02439
Detection Prevalence    0.9555 1.961e-04   0.04435
Balanced Accuracy       0.5720 5.011e-01   0.57693
```

\- Notes (Will be expanded and explained in future update):

-   Accuracy: 85.09%

-   Target variable may not be balanced

#### Random Forest - All Features

```         
Confusion Matrix and Statistics

          Reference
Prediction    X0    X1    X2
        X0 21555   431  3514
        X1     0     0     0
        X2     0     0     0

Overall Statistics

               Accuracy : 0.8453          
                 95% CI : (0.8408, 0.8497)
    No Information Rate : 0.8453          
    P-Value [Acc > NIR] : 0.5042

                  Kappa : 0

 Mcnemar's Test P-Value : NA

Statistics by Class:

                     Class: X0 Class: X1 Class: X2
Sensitivity             1.0000    0.0000    0.0000
Specificity             0.0000    1.0000    1.0000
Pos Pred Value          0.8453       NaN       NaN
Neg Pred Value             NaN    0.9831    0.8622
Prevalence              0.8453    0.0169    0.1378
Detection Rate          0.8453    0.0000    0.0000
Detection Prevalence    1.0000    0.0000    0.0000
Balanced Accuracy       0.5000    0.5000    0.5000
```

\- Notes (Will be expanded and explained in future update):

-   Model is only predicting '0', likely an error in model setup

-   Due to inaccurate predictions, accuracy is not applicable.

#### KNN - All Features

-   in progress

## Full analysis

***\*To be updated in later portion of project***

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

***\*To be updated in later portion of project***

## Strengths and Limitations

***\*To be updated in later portion of project***

-   LIMIT: Having a lot of categorical features

-   LIMIT: Having a response variable with 3 levels instead of a binary.

## Conclusions

***\*To be updated in later portion of project***

{{< pagebreak >}}

# References

***\*To be updated in later portion of project***